---
title: "CAB220 Portfolio 2"
author: "KA LONG LEE (N9845097)"
date: "28/09/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CAB220 Portfolio2

Overview
This portfolio accounts for 20% of your overall grade of CAB220. Full mark of this portfolio is 20. The tasks in this portfolio are designed to assess your knowledge and skills in

• Descriptive statistical data analysis and visualisation 
• Statistical hypothesis testing
• Linear regression
• Logistic regression

Data:

The fictitious data set for this portfolio includes the records of 2,550 first-year students of an Australian university in terms of case ID, Attrition, Degree Type, Achieved Credit Points, Attendance Type, Age, Failed Credit Points, International student, First in family in university, Gender, GPA, OP Score, Socio Economic Status, Teaching Period Admitted,and Faculty.

Working Environment Configuration:

```{r cars}

# Import Library
library(ggplot2)
library(dplyr)

# Data Preprocessing Library
library(caTools)

# Logistic Regression Library
library(DAAG)
library(cowplot)
library(InformationValue)

# Setting up the working directory
# So that It can import external file
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# Import external files
# Most of the visualization function is stored in this file
# Please check it if you are interested in the code
source("data_visualization.R")

# Import Data 
uniData <- read.csv("datasets/Portfolio_2_data.csv", header = TRUE) %>%
  select(2:15)

```

## Task 1 Summarise the information in each variable (except case ID) using a table or an appropriate statistical graph

#### Summary each variables using a table

```{r echo=FALSE}
uniData %>%
  summary()
```
#### Summary each categorical data in uni dataframe using appropriate graphs

```{r}
visualize_categorical_data(uniData)
```

#### Summary each numerical data in uni dataframe using appropriate graphs

```{r}
# A function print out each appropriate graphs
visualize_numerical_data(uniData)
```

## Task 2 Compare average GPA between male and female students using a graph, conduct a statistical test, and interpret its results

```{r}
# Compare average GPA between Male and Female 
# Conduct a statistical Test
# Interpret its results
visualize_boxplot_gpa_vs_gender(uniData)
```


### T-Test & Variance
```{r}
# T Test
t.test(uniData$GPA ~ uniData$Gender)
# Variance
var.test(uniData$GPA ~ uniData$Gender)
```


## Task3 Explore the relationship between OP Score and GPA using a graph, describe the relationship
```{r}
visualize_relationship_op_and_gpa(uniData)
```


## Task 4 Linear Regression

Develop a linear regression model of GPA using the given data. You need to describe your choice of predictors, examine your model’s assumptions, assess model fit, and interpret the final model’s regression coefficients.

Analyse Each numbemerical data its relation related to GPA

```{r}
visualize_scatterplots_Vs_GPA(uniData)

```

### Spiting dataframe into training set and test set
```{r}
set.seed(2)

split <- sample.split(uniData, SplitRatio = 0.7)
train <- subset(uniData, split==TRUE)
test <- subset(uniData, split==FALSE)

```

### Training Linear Regression Model
```{r}
linear_model <- lm(GPA ~ Achieved_Credit_Points, data=train)
summary(linear_model)
```
### Plot the Linear Regression Prediction Line
```{r}

plot(linear_model)

```

## Task 5 Logistic Regression